Refining from the BON High-level Specification Language to the VHDL-AMS Hardware Description Language

### 2009-08-31 papers/BON_VHDL/paper.tex
added more

\documentclass{article}
\usepackage{times}
\usepackage{multirow}
\usepackage{ifpdf}
\usepackage{a4wide}
\usepackage{pdfsync}
\usepackage{amsmath}
\usepackage{listings}

@@

\ifpdf
\usepackage[pdftex,bookmarks=false,a4paper=false,
            plainpages=false,naturalnames=true,
            colorlinks=true,pdfstartview=FitV,
            linkcolor=blue,citecolor=blue,urlcolor=blue,
            pdfauthor="Joseph R. Kiniry and Dragan Stosic"]{hyperref}
\else
\usepackage[dvips,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
\fi

@@

@@

%======================================================================
\section{Introduction}
\label{sec:introduction}
The \ST sensor platform, represents heterogeneous collection of digital, analog
and mixed signal hardware circuits. Embedded system complexity
requires the need to strive for formal descriptions each component
separately in order to produce specification of the \STS.
Our aim is to be able to precisely describe the structures of circuits,
the behavior of circuits and the formal link between structures and behaviors.
In other words, the behavioral specification of the \STS may be
represented as a \inv on the signals at the circuit's ports and may be obtained
by taking the conjunction of the \emph{invariants} that describe
behaviors of the individual components separately.
For this purpose we need  sophisticated tool and methodology
in order to be able to achieve this goal. 
The authors will illustrate in this article how BON can represent a high
level analogue mapping and will further illustrate a procedure for
transformation from VHDL-AMS to BON.\\
This article provides two main contributions.
\begin{enumerate}
\item
 It provides a solution how to formally emulate
 piece of hardware using BON-JML approach; up to date BON and JML
 have been used mostly in the pure object oriented software construction.   
\item
It relates hardware description code and formal calculus.
This means that if there  is a code that is "correct by construction",
then it will be derivable from BON (informal/formal) specification,
via the refinement calculus to JML implementation.    
\end{enumerate}
The rest of this article is organized as follows.
In \autoref{sec:vhdlams} we have  presented an overview of VHDL-AMS language.
\autoref{sec:choiceobon} defines  BON principles and explains the
reasons why we have decided to use BON method.
The central part of the our article is \autoref{sec:mappingVB}, which
shows mapping between VHDL-AMS and BON.
In \autoref{sec:refVB} we will give practical aspect of refinement,
applied on real capacitor component, in order to produce formal specification 
and software simulation.\\ 
A comprehensive list of the mapping between BON and VHDL-AMS can be
found in the appendix.
%=====================================================================

@@

\begin{itemize}
\item
\emph{Terminals} represents phisical connection points or circuit nodes
of a network connectivity. They are declared to be of various
\emph{natures}, which represent different energy domains of a system.
Terminals hold no values; their only role is to facilitate the formation
of conservative equation sets that, in turn, constrain the associated
\emph{branch quatities}.
\item
The \emph{quantities} denote a waveform or a time series of values.
The behaviour of the \emph{quantities} is given by the \emph{simultaneous
 statements}. They take values as a result of solving the set of simultaneous
ordinary differential and algebraic equations (DAEs). To simplify,
quantities can be understood as analog unknowns of (DAEs) that specify
analog behaviour. There are three kinds of quantity in VHDL-AMS: free,
branch and source quantities. A \emph{free quantity} can be used in
signal-flow modeling. A \emph{branch quantity} is used to model conservative energy
systems. A \emph{source quantity} is used for frequency and noise modeling.
\item
The \emph{simultaneous statements} creates one or more characteristic
equations that are solved by the dedicated simulation kernel a.k.a.
"analog solver", to model continuously valued behaviour. 
\item
The \emph{procedural statements} or \emph{procedural} express analog behaviour
using sequential statements. Actually procedural describes analog behaviour
by computing new values for quantities from the current value of quantities.
\item
The \emph{break statement} is an construct which allows to
initialize the DAE system at the beginning of a transient
simulation and to mark discontinuities during such a transient
simulation. A break statement allows restart of the analog kernel,
than overrides quantity values and executes concurrently with the analogue model.
For any concurrent break statement there is an associated process.
\end{itemize}

\section{Choice of BON as the modeling tool}
\label{sec:choiceobon}
%Brief description and motivation.
The Business Object Notation (BON) is an analysis and design notation
for object-oriented systems  which emphasizes \emph{seamlessness},\xspace
\emph{reversibility} and software \emph{contracting}.
Let us define these principles precisely and point out why the principles
are important in VHDL-AMS simulation.

@@

different levels of abstraction. For example, one may wish to analyze

@@

\emph{quantities}. When we reasoning about conservation of energy law, for example,

@@

However if our implementation does not satisfy conservation of energy law, 

@@

In practice, one can use \emph{invariant} against the already existing code in
order to check the correctness.
It is important because we cannot patch a fabricated chip once company started

@@

An VHDL-AMS design units are described using an \ent-\arch structure.
The design unit contains an \ent and one or more \archs.
The \ent specifies the interface that an entity provides of itself
to the outside. In other words, \ent introduces a name, list of
input/output ports and the definition of the generic parameters.
The textual presentation of \ent is shown in Figure.1.
\input{adc-entity}

The \arch body represents description of the internal implementation
of an \ent including only process statements(encapsulation), which
are collections  of actions to be executed in sequence, or simultaneous statements,
which describe equations of analogue system behaviour.
The textual presentation of \arch is shown in Figure.2.
\input{adc-arch}

>>>>

\documentclass{article}
\usepackage{times}
\usepackage{multirow}
\usepackage{ifpdf}
\usepackage{a4wide}
%\usepackage{pdfsync}
\usepackage{amsmath}
\usepackage{listings}

@@

\ifpdf
\usepackage[pdftex,bookmarks=false,a4paper=false,
            plainpages=false,naturalnames=true,
            colorlinks=true,pdfstartview=FitV,
            linkcolor=blue,citecolor=blue,urlcolor=blue,
            pdfauthor="Joseph R. Kiniry and Dragan Stosic"]{hyperref}
\else
\usepackage[dvips,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
\fi
\ifpdf
  \DeclareGraphicsRule{*}{mps}{*}{}
\fi

@@

\def\sectionautorefname{Section}

@@

%======================================================================
\section{Introduction}
\label{sec:introduction}
The \ST sensor platform, represents heterogeneous collection of digital, analog
and mixed signal hardware circuits. Embedded system complexity
requires the need to strive for formal descriptions each component
separately in order to produce specification of the \STS.
Our aim is to be able to precisely describe the structures of circuits,
the behavior of circuits and the formal link between structures and behaviors.
In other words, the behavioral specification of the \STS may be
represented as a \inv on the signals at the circuit's ports and may be obtained
by taking the conjunction of the \emph{invariants} that describe
behaviors of the individual components separately.
For this purpose we need  sophisticated tool and methodology
in order to be able to achieve this goal. 
The authors will illustrate in this article how BON can represent a high
level analogue mapping and will further illustrate a procedure for
transformation from VHDL-AMS to BON.\\
This article provides three main contributions.
\begin{enumerate}
\item
 It provides a solution how to formally emulate
 piece of hardware using BON-JML approach; up to date BON and JML
 have been used mostly in the pure object oriented software construction.   
\item
It relates hardware description code and formal calculus.
This means that if there  is a code that is "correct by construction",
then it will be derivable from BON (informal/formal) specification,
via the refinement calculus to JML implementation.
\end{enumerate}
The rest of this article is organized as follows.
In \autoref{sec:vhdlams} we have  presented an overview of VHDL-AMS language.
\autoref{sec:choiceobon} defines  BON principles and explains the
reasons why we have decided to use BON method.
In \autoref{sec:mappingVB} we will show mapping between VHDL-AMS and BON.
A comprehensive list of the mapping between BON and VHDL-AMS can be
found in the appendix. 
The central part of the our article is \autoref{sec:refVB}, in which
we will give practical aspect of refinement, applied on lowpass filter, 
in order to produce formal specification and software simulation.\\ 
%=====================================================================

@@

\begin{itemize}
\item
\emph{Terminals} represents phisical connection points or circuit nodes
of a network connectivity. They are declared to be of various
\emph{natures}, which represent different energy domains of a system.
Terminals hold no values; their only role is to facilitate the formation
of conservative equation sets that, in turn, constrain the associated
\emph{branch quatities}.
\item
The \emph{quantities} denote a waveform or a time series of values.
The behaviour of the \emph{quantities} is given by the \emph{simultaneous
 statements}. They take values as a result of solving the set of simultaneous
ordinary differential and algebraic equations (DAEs). To simplify,
quantities can be understood as analog unknowns of (DAEs) that specify
analog behaviour. There are three kinds of quantity in VHDL-AMS: free,
branch and source quantities. A \emph{free quantity} can be used in
signal-flow modeling. A \emph{branch quantity} is used to model conservative energy
systems. A \emph{source quantity} is used for frequency and noise modeling.
\item
The \emph{simultaneous statements} creates one or more characteristic
equations that are solved by the dedicated simulation kernel a.k.a.
"analog solver", to model continuously valued behaviour. 
\item
The \emph{procedural statements} or \emph{procedural} express analog behaviour
using sequential statements. Actually procedural describes analog behaviour
by computing new values for quantities from the current value of quantities.
\item
The \emph{break statement} is an construct which allows to
initialize the DAE system at the beginning of a transient
simulation and to mark discontinuities during such a transient
simulation.For any concurrent break statement there is an associated process.
A break statement allows restart of the analog kernel,
than overrides quantity values and executes concurrently with the analogue model.

\end{itemize}

\section{Choice of BON as the modeling tool}
\label{sec:choiceobon}
%Brief description and motivation.
The Business Object Notation (BON) is an analysis and design notation
for object-oriented systems  which emphasizes \emph{seamlessness},\xspace
\emph{reversibility} and software \emph{contracting}.
Let us define these principles precisely and point out why the principles
are important in VHDL-AMS simulation.\\

@@

different levels of abstraction. For instance, one may wish to analyze

@@

\emph{quantities}. For instance, when we reasoning about conservation of energy law,

@@

However, if our implementation does not satisfy conservation of energy law, 

@@

Another conclusion may be deducted from the example given above.\\
Intuitively \inv may be useful for testing the function of the hardware components.
In practice, one can runs simulator against the component in order to check 
the properties. Testing can then be done by checking whether the \inv hold or not.
For example, the most common problem of the real components is the glich - an electrical pulse 
of short duration that is usually the result of a design error.
The number of glitches that occur can be known by the fact how many times per day, 
the \inv was broken and reestablished. 
It is important because we cannot patch a fabricated component once we started

@@

\includegraphics{circ.1}
